---
title: "aggregate-spatial-data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{aggregate-spatial-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Besides building grids, this package includes functions for aggregating spatial
data. To start, let's load some packages:

```{r setup}
library(sf)
library(terra)
library(treeslabgrid)
```

Now let's get some raster data and throw a grid over it:

```{r get_data}
r <- rast(system.file("ex/meuse.tif", package = "terra"))
g <- make_grid_origin_res(
  xy_origin = get_center(r),
  xy_min = get_min(r),
  xy_max = get_max(r),
  cell_size = 500,
  crs = terra::crs(r),
  id_col = "grid_id"
)
```
Note the use of the functions `get_center`, `get_min`, and `get_max` to get the
extent coordinates of objects (`sf` or `terra`).

Now plot, just to make sure the grid makes sense:

```{r plot_r_grid}
plot(r, reset = FALSE)
plot(st_cast(g[get_geom_colname(g)], "LINESTRING"), add = TRUE)
```

Now we can use this grid to aggregate the data it covers:

```{r aggregate_meuse}
meuse_agg <- aggregate_raster(
  x = r,
  by = NA_character_,
  grid = g,
  grid_id = "grid_id",
  funs = c("sum", "min", "mean", "max"),
  na.rm = TRUE
)
```

Just a note about how the column names are built in the results:

```{r aggregate_meuse_coords}
colnames(meuse_agg)
```

Besides the identifier of each cell in the grid (`grid_id`), we have columns
for the combination of each aggregation function (`funs`), the input raster
(`meuse`), and each category (we didn't include one in our example).
The names of the column in the results are build in the same order, separated
by points.
The exception is `n`, which correspond to the number of raster cells included
in each cell of the aggregation grid.

Now, we can, for example, plot the total value in each cell of the aggregation
grid:

```{r aggregate_meuse_plot_sum}
plot(meuse_agg["sum.meuse.1"])
```

Or the mean value:

```{r aggregate_meuse_plot_mean}
plot(meuse_agg["mean.meuse.1"])
```
